---
title: "04_3_BayesLogReg"
author: "Kyla McConnell"
date: "2/9/2022"
output: html_document
---

# Bayesian Logistic Regression
```{r setup, include=FALSE}
library(tidyverse) 
library(brms)

# data subsetted from Oberauer 2019
df_recall_data <- read_csv("PairsRSS1_all.csv") %>%
    mutate(correct = if_else(response_category ==1, 1, 0)) %>%
    filter(response_size_list + response_size_new_words == 0) %>%
    filter(subject == 10) %>%
    mutate(c_set_size = set_size - mean(set_size)) %>%
    select(subject, set_size, c_set_size, correct, trial)
```

Data: working memory -- one participant recalls a prompted word from a set of size 2, 4, 6, or 8 (centered)

# Bernoulli distribution

- the Bernoulli likelihood generates a  0 or 1 response with a particular probability, theta (θ)
- the output is the probability of success 
- there is no residual error term

```{r}
as.numeric(rbernoulli(n=10))
```

- theta must be bound between 0 and 1 because it is a probability --> so the Gaussian/normal likelihood wouldn't make sense (it allows any numeric value, -infinity to infinity)
- Logit link converts probability of getting a 1 to a numeric range from -infinity to infinity using log odds

- log odds vs. probability: https://towardsdatascience.com/https-towardsdatascience-com-what-and-why-of-log-odds-64ba988bf704

- log odds convert probabilities (which are between 0 and 1) to the full range of numbers -infinity to infinity
- the logit link is the inverse of the logistic link (so they are different things! and one undoes the other, like multiplication and division)

- Fazit: the probability of success is calculated on the log-odds scale via the logit link because this allows you to link a probability to the full range of real numbers, then converted back to probability space using the logistic link (the opposite of the logit link)

## alpha
- alpha (𝛼) is the intercept -- the log odds of success at the average value of your predictors (given that they are centered, otherwise it would be when your predictors are at 0)
  - in this example: log odds of successfully recalling the word at the average set size of 5 words
- an example prior with maximal uncertainty might be that there's a 50/50 chance of success
- alpha is not described in terms of probability though, it's log odds
- but! be careful with the SD because log odds and probability are not linearly related. if you have a large SD, it will put a lot of weight at the tails, i.e. at 0 or 1 probability
- you can use the qlogis and glogis functions to translate log odds to probability    for more intuitive understanding
  
## beta
- beta is the change in log odds when you change the independent variable, so here, by increasing the set size 
- bruno changes his prior_pred_fast function to work in the logistic case
  
```{r}
logistic_model_pred <- function(alpha_samples, beta_samples, set_size, N_obs) { 
  map2_dfr(alpha_samples, beta_samples,
            function(alpha, beta) { 
              tibble(set_size = set_size,
                      c_set_size = set_size - mean(set_size), 
                      theta = plogis(alpha + c_set_size * beta),
                      correct_pred = rbernoulli(N_obs, p = theta)) }, .id = "iter") %>%
            mutate(iter = as.numeric(iter))
}
```

- Try out four different standard deviations: 1, .5, .1, .01, .001 on a sample of 800 observations of 4 different set sizes (increased N so that it's easier to see what the prediction would be in a larger sample)
```{r}
N_obs <- 800
set_size <- rep(c(2, 4, 6, 8), 200)

alpha_samples <- rnorm(1000, 0, 1.5) # assume this is the alpha we've decided on
sds_beta <- c(1, 0.5, 0.1,0.01, 0.001) # these are the SDs for beta that we're trying out

prior_pred <- map_dfr(sds_beta, function(sd) {
  beta_samples <- rnorm(1000, 0, sd) # center the normal dist for beta at 0 and just try different SDs
  logistic_model_pred(alpha_samples = alpha_samples,
                      beta_samples = beta_samples, 
                      set_size = set_size,
                      N_obs = N_obs) %>%
  mutate(prior_beta_sd = sd) 
})
```


Take a look at each of the priors we tried and plot:
```{r}
(mean_accuracy <- prior_pred %>%
group_by(prior_beta_sd, iter, set_size) %>% 
  summarize(accuracy = mean(correct_pred)) %>% 
  mutate(prior = paste0("Normal(0, ",prior_beta_sd,")")))
```

```{r}
mean_accuracy %>% 
  ggplot(aes(accuracy)) + 
  geom_histogram() + 
  facet_grid(set_size ~ prior)
```

- Normal(0, 1) doesn't look good because there is a lot of weight in the 0 and 1 in the probability space 
- You can see the same basic structure in Normal(0, .5)
- The other 3 look very similar, so we can look at the difference when comparing different set sizes (i.e. 2 to 4, 4 to 6 and 6 to 8)

```{r}
(diff_accuracy <- mean_accuracy %>%
arrange(set_size) %>%
group_by(iter, prior_beta_sd) %>%
mutate(diffaccuracy = accuracy - lag(accuracy) ) %>% mutate(diffsize = paste(set_size,"-", lag(set_size))) %>% filter(set_size >2))

diff_accuracy %>% 
  ggplot(aes(diffaccuracy)) + 
  geom_histogram() + 
  facet_grid(diffsize ~ prior)
```

- Bruno selects Normal(0, 0.1) because it has the most uncertainty without being unreasonable


??? "In principle, we could produce the prior predictive distributions using brms with sample_prior = "only" and then predict()"

# Fit logistic regression
- Be sure to define link = logit (there are others, think this is the default)

```{r}
fit_recall <- brm(correct ~ 1 + c_set_size, data = df_recall_data,
family = bernoulli(link = logit),
prior = c(
prior(normal(0, 1.5), class = Intercept),
prior(normal(0, .1), class = b, coef = c_set_size) )
)
```

The output is in log-odds 
```{r}
posterior_summary(fit_recall, pars = c("b_Intercept", "b_c_set_size"))
```

- The beta for set size shows us that accuracy is reduced as set size increases

```{r}
plot(fit_recall)
```

- Chains are mixing so it looks good

## Report the model output

If we want to talk about the effect estimated by the model in log-odds space, we summarize the posterior of 𝛽 in the following way:𝛽̂=−0.18,95%CrI=[−0.34,−0.03].

Can also translate this back to probability space, first with the intercept/alpha
```{r}
alpha_samples <- posterior_samples(fit_recall)$b_Intercept 
av_accuracy <- plogis(alpha_samples)

c(mean = mean(av_accuracy), quantile(av_accuracy, c(.025, .975)))
```

But also with beta for set size, i.e. the effect of the change in set size
```{r}
beta_samples <- posterior_samples(fit_recall)$b_c_set_size
effect_av_set_size <- plogis(alpha_samples) - plogis(alpha_samples - beta_samples) 

c(mean = mean(effect_av_set_size), quantile(effect_av_set_size, c(.025, .975)))
```

Interpretation: if we increase the set size from the average set size minus one to the average set size (5), we get a reduction in the accuracy of recall of −0.02, 95% CrI = [−0.04, 0].
- -.02 here means a decrease of 2%

But since a set size of 5 didn't actually exist, we could check more meaningful numbers
```{r}
set4 <- 4 - mean(df_recall_data$set_size)
set2 <- 2 - mean(df_recall_data$set_size)
effect_4m2 <- plogis(alpha_samples + set4 * beta_samples) - plogis(alpha_samples + set2 * beta_samples)

c(mean = mean(effect_4m2), quantile(effect_4m2, c(.025, .975)))
```

