---
title: "Quarto demonstration"
description: 
date: today 
author:
  - name: 
    url: 
    affiliation: 
    affiliation-url: 
title-block-banner: true
format:
  html:
    theme: 
---


# Preparation

## Choosing the data

As everyone always says,

penguins are very cute!

...so let's work with penguin data today.  

Note that the palmerpenguins dataset is meant as an alternative to the commonly used iris data.

More information on this data set is available on Allison Horst's Github page: https://allisonhorst.github.io/palmerpenguins/


Data were collected and made available by Dr. Kristen Gorman and the Palmer Station, Antarctica LTER, a member of the Long Term Ecological Research Network.


This data can also be used via the {palmerpenguins} package and was originally introduced in Gorman and colleagues' article. 


## Loading packages & reading data

First, we need to:  
load **packages**  
read in **data**  
check if that data is *in the right format*  

By "right format", we mean that the data is tidy, and ready to be summarised and graphed.

Using the data without downloading it
This same dataset is also available in the package palmerpenguins.

```{r library and data}
#| warning: false
#| message: false
library(tidyverse)
penguins <- read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-07-28/penguins.csv') %>% 
  drop_na()
```

```{r inspection}
head(penguins)
```

We've removed missing values here, which means that the data has `r nrow(penguins)` rows now.


## Some descriptive statistics

The average bill length is ? mm and the average bill depth is ? mm.

The data was collected between ? and ?.


### Comparing base R and tidyverse syntax

Here's code to see data in a specific column:

#### base R
```{r}
penguins$bill_length_mm
```

#### tidyverse
```{r}
penguins %>% select(bill_length_mm)
```


# Graphs

Some ideas for graphs:

penguin weight by flipper length  
for the entire data  
or separately for each species  
and additionally by sex  
flipper and bill length  
bill length and depth  


## Weight and flipper length

The following graphs are slightly simplified versions from the vignette for the {palmerpenguins} package: https://allisonhorst.github.io/palmerpenguins/articles/examples.html


### Entire data
```{r}
mass_flipper <- ggplot(data = penguins,
                       aes(x = flipper_length_mm,
                           y = body_mass_g)) +
  geom_point(size = 3,
             alpha = 0.8) +
  theme_minimal() +
  labs(title = "Penguin size, Palmer Station LTER",
       subtitle = "Flipper length and body mass for Adelie, Chinstrap and Gentoo Penguins",
       x = "Flipper length (mm)",
       y = "Body mass (g)") +
  theme_minimal()

mass_flipper
```


### Separated by species
```{r}
mass_flipper_species <- ggplot(data = penguins,
                       aes(x = flipper_length_mm,
                           y = body_mass_g)) +
  geom_point(aes(color = species,
                 shape = species),
             size = 3,
             alpha = 0.8) +
  theme_minimal() +
  scale_color_manual(values = c("darkorange","purple","cyan4")) +
  labs(title = "Penguin size, Palmer Station LTER",
       subtitle = "Flipper length and body mass for Adelie, Chinstrap and Gentoo Penguins",
       x = "Flipper length (mm)",
       y = "Body mass (g)",
       color = "Penguin species",
       shape = "Penguin species") +
  theme_minimal()

mass_flipper_species
```

### Separated by species and sex
```{r}
mass_flipper_species +
  facet_wrap(~sex)
```

